---
title: "Nonnamaker Exercise 10"
output: html_document
---

## Question 1. 

```{r include=FALSE}
rm(list=ls())
setwd("/Users/emilynonnamaker/Box/personalStuff/School/PhD/biocomputingFall2018/Exercises/Biocomp-Fall2018-181109-Exercise10")
dat <- read.csv("data.txt")
library(ggplot2)
library(arm)
```

hump-shaped vs. linear function of x by comparing a quadratic
(a + bx + cx2) vs. linear (a + bx) model.
Generate a script that evaluates which model is more appropriate for the data in data.txt.

```{r}
N=length(dat)
x=dat$x
y=dat$y
# plot our observations
ggplot(dat,aes(x=x,y=y))+geom_point()+theme_classic() # nice line. 
```

For the first model (linear)
```{r}
linlike<-function(p,x,y){
B0=p[1]
B1=p[2]
sigma=exp(p[3])
expected=B0+B1*x
linll=-sum(dnorm(x=y,mean=expected,sd=sigma,log=TRUE))
return(linll)
}
```

```{r}
initialGuess=c(1,1,1) # for B0, B1, and sigma
fitlin=optim(par=initialGuess,fn=linlike,x=x,y=y)
# fit is a variable that contains a list describing the result of minimization
print(fitlin) # it converged, good

lm<-lm(y ~ x) 
display(lm) # sanity check
```

For the second model (quadratic)

```{r}
quadlike<-function(p,x,y){
B0=p[1]
B1=p[2]
C=p[3]
sigma=exp(p[4])
expected=B0+B1*x+C*I(x^2)
quadll=-sum(dnorm(x=y,mean=expected,sd=sigma,log=TRUE))
return(quadll)
}
```

```{r}
initialGuess=c(1,1,1,1) # for B0, B1, C, and sigma.
fitquad=optim(par=initialGuess,fn=quadlike,x=x,y=y)
# fit is a variable that contains a list describing the result of minimization
print(fitquad) # also converged, nice
```

To test the two:

```{r}
t.stat=2*(fitlin$value-fitquad$value)
t.stat

df=length(fitquad$par)-length(fitlin$par)
df

1-pchisq(t.stat, df=1)
```

Plot it up to check it out:

```{r, warning=FALSE}
plot(y~x)
abline(lm, col='red')
abline(lm(y~x+I(x^2)))
```

Looking at p-values, there's no real difference between the model fits (this can be seen when plotting as well), so then by parsimony, the linear model is best. 

## Question 2

A classic model of competition between two species was developed by Lotka & Volterra. This model has two state variables described by two differential equations:

* $dN1/dt = R1(1 − N1α11 − N2α12)N1$

* $dN2/dt = R2(1 − N2α22 − N1α21)N2$

The criteria for coexistence of two species in the Lotka-Volterra competition model is α12 < α11 and α21 < α22

Generate a script that uses three or more model simulations to demonstrate the validity of these criteria for coexistence.

```{r include=FALSE}
library(deSolve)
```

Model: 

```{r}
LVSim<-function(t,y,p){
  N1=y[1]
  N2=y[2]
  R1=p[1]
  R2=p[2]
  a11=p[3]
  a12=p[4]
  a21=p[5]
  a22=p[6]
  
  dN1dt=R1*(1-(N1*a11)-(N2*a12))*N1
  dN2dt=R2*(1-(N2*a22)-(N1*a21))*N2

  return(list(c(dN1dt,dN2dt)))
}
```

# Case 1

Correct (for my own refrence) 
α11 > α12 and α21 < α22
(big), (small), (small), (big)

First half is correct
α11 > α12 and α21 > α22

```{r}
times=1:100
y0=c(0.1,0.1)
#params2=c(R1, R2, a11 (big), a12 (small), a21(big), a22(small))
params2=c(0.5, 1, 2, 1, 2, 1)
sim2=ode(y=y0,times=times,func=LVSim,parms=params2)
out2=data.frame(time=sim2[,1],pop=sim2[,2],case=sim2[,3])
ggplot(out2,aes(x=time,y=pop))+geom_line()+geom_line(data=out2,mapping=aes(x=time,y=case),col='red')+theme_classic()
```

One population dominates the other. 

# Case 2

Correct (for my own refrence) 
α11 > α12 and α21 < α22
(big), (small), (small), (big)

Second half is correct
α11 < α12 and α21 < α22

```{r}
times=1:100
y0=c(0.1,0.1)
#params3=c(R1, R2, a11 (small), a12 (big), a21(small), a22(big))
params3=c(0.5, 1, 1, 2, 1, 2)
sim3=ode(y=y0,times=times,func=LVSim,parms=params3)
out3=data.frame(time=sim3[,1],pop=sim3[,2],case=sim3[,3])
ggplot(out3,aes(x=time,y=pop))+geom_line()+geom_line(data=out3,mapping=aes(x=time,y=case),col='red')+theme_classic()
```

Again, one population dominates the other. 

# Case 3

Correct (for my own refrence) 
α11 > α12 and α21 < α22
(big), (small), (small), (big)

Both correct
α11 > α12 and α21 < α22

```{r}
times=1:100
y0=c(0.1,0.1)
#params4=c(R1, R2, a11 (big), a12 (small), a21(small), a22(big))
params4=c(0.5, 1, 2, 1, 1, 2)
sim4=ode(y=y0,times=times,func=LVSim,parms=params4)
out4=data.frame(time=sim4[,1],pop=sim4[,2],case=sim4[,3])
ggplot(out4,aes(x=time,y=pop))+geom_line()+geom_line(data=out4,mapping=aes(x=time,y=case),col='red')+theme_classic()
```

The populations end up equalizing, and co-existing. 

# Case 4

Correct (for my own refrence) 
α11 > α12 and α21 < α22
(big), (small), (small), (big)

None correct
α11 < α12 and α21 > α22

```{r}
times=1:100
y0=c(0.1,0.1)
params5=c(0.5,1,1,2,2,1)
sim4=ode(y=y0,times=times,func=LVSim,parms=params5)
out4=data.frame(time=sim4[,1],pop=sim4[,2],tumor=sim4[,3])
ggplot(out4,aes(x=time,y=pop))+geom_line()+geom_line(data=out4,mapping=aes(x=time,y=tumor),col='red')+theme_classic()
```

One population completely oucompetes the other. 